The Tor Network - Strengths and Weaknesses \\ \Large Computer Security 1 - CS5460 \\ \LaTeX{}

### 2013-11-29 paper.tex
trimming of cipher descriptions

RSA is an algorithm for public-key cryptography. Its security is based on the infeasibility of factoring the product of two large primes. RSA, like all other public-key cryptography algorithms, relies upon two keys: one public and the other private. The public key may be published and is used for encryption and for verifying digital signatures. The private key is used for decryption, the generation of digital signatures. Thus, through RSA, only the owner of the private key can decrypt incoming messages, and only they can digitally sign outbound messages. This makes RSA extremely powerful for both authenticity and privacy.

To generate the RSA keys, two distinct prime numbers of similar bit-length are chosen at random, \textit{p} and \textit{q}. \textit{n} is computed as the product of \textit{p} and \textit{q}, and it is used as the modulus for both the public and private keys. The bit-length of \textit{n} is the RSA key length. Then let $ \varphi(n) = \varphi(p)\varphi(q) = (p - 1)(q - 1) $ where $ \varphi $ is Euler's totient function, which counts the number of positive integers less than or equal to \textit{n} that are relatively prime to \textit{n}. Then \textit{e} is chosen such that \textit{e} and $ \varphi(n) $ are coprime. \textit{d} is also chosen such that it is the multiplicative inverse of e modulo $ \varphi(n) $. The RSA public key then consists of the modulus \textit{n} and the exponent \textit{e}. The private key likewise consists of \textit{n} and the exponent \textit{d}.

A message \textit{M} is encrypted by first blocks of length less than \textit{n} using an agreed-upon padding scheme. The ciphertext \textit{c} is then computed as $ m^e \bmod n $, which can be calculated quickly using the method of exponentiation by squaring. The owner of the private key can then recover the plaintext \textit{M} by computing $ c^d \bmod n $, and then reversing the padding scheme. The decryption process is often accelerated by optimizations such as the Chinese Remainder Theorem.

Furthermore, to sign a message, the owner of the private key first uses an agreed-upon hash function to compute a hash value of the message, raises it to the power of \textit{d} modulo \textit{n} and then sends that in conjuction with \textit{M}. The recipient, who only has access to the public keys, then raises the signature to the power of \textit{e} modulo \textit{n} and compares the result with his hash of the message using the same hash algorithm. If the two match, then he can prove the authenticity and integrity of the message. These two capabilities make RSA an extremely powerful tool for securing authenticity, integrity, and privacy of communication between two parties.

@@

Elliptic curves over the real number system can be defined through the equation, $ y^2 = x^3 + ax + b $ where \textit{a} and \textit{b} are real numbers. This formula describes a plane curve through the Euclidean plane, although elliptic curves do exists in other spaces. However, current cryptographic purposes only work with elliptic curves within the context of the real number systems.

A smaller key size is one of the most significant benefits introduced by ECC. Current NIST recommendations state that a key size of 160 bits for ECC offers a similar level of security to 1024 bits of RSA/DH. Likewise, 224 bits for ECC is analogous to 2048 bits of RSA/DH. It should be noted that the required key sizes for ECC increases significantly slower than the key sizes for RSA and Diffie-Hellman increase at a much faster rate than the required key size for ECC. Thus ECC offers more security per increase in key size than for RSA. The smaller key size also directly correlates to an increase in computational efficiency; as the bit size increases, the speed difference between Diffie-Hellman and ECC grows superlinearly.\cite{CaseForECC}

@@

Once the SSL/TLS handshake completes, both parties use the generated session key and turn to symmetric-key encryption for hiding of data.

Data Encryption Standard (DES) is a block cipher introduced in 1977 by IBM. It was considered insecure in 1998 when the DES Challenges demonstrated that it was feasible to brute-force DES ciphertexts in a matter of hours using custom hardware. This weakness was partially mitigated by the introduction of triple DES (3DES) in the same year, where three rounds of DES were performed using three separate 56-bit keys. However, 3DES was found to be susceptible to meet-in-the-middle, chosen-plaintext, and known-plaintext attacks, so despite using a 168-bit key, 3DES only provides 80 bits of security.\cite{NIST2007} 3DES is still used on the web and by older Microsoft products, but has been largely superseded by other algorithms.

Advanced Encryption Standard (AES) is a block cipher introduced in 2001. It supersedes 3DES, and is now used extensively worldwide. AES became the recommended cipher for protecting classified documents for the US government in 2003, and side-channel attacks on specific implementations were the only known vulnerabilities up until mid 2009. To date, the most successful public attacks require $ 2^126.1 $ operations to recover a 128-bit key,\cite{Nikolic2009} and related-key attacks can break AES-192 and AES-256 in $ 2^176 $ and $ 2^99.5 $ operations,\cite{Bogdanov2011} respectively.

RC4 is a stream cipher invented in 1987 and sunsequently leaked in 1994. The RC4 algorithm is simple and extremely fast in software, and is now the most widely used software stream cipher, most notably in protocols such as TLS and WEP. Although it is popular, a number of practical attacks and weaknesses have been demonstrated against its Key Scheduling Algorithm and PRNG that call into question RC4's security. In November 2013, Microsoft recommended disabling RC4 from TLS wherever possible.

Robert Graham's analysis of the incoming handshakes to his exit node revealed that almost 90\% of the circuits agreed on the AES cipher, while the remainder used 3DES. For communication to web servers, RC4 remained the most popular algorithm.

@@

>>>>

RSA is an algorithm for public-key cryptography. Its security is based on the infeasibility of factoring the product of two large primes. RSA, like all other public-key cryptography algorithms, relies upon two keys: one public and the other private. The public key may be published and is used for encryption and for the verification of digital signatures. The private key is used for decryption and the generation of digital signatures. Thus, only the owner of the private key can decrypt incoming messages, and only they can digitally sign outbound messages. These properties make RSA and other public-key algorithms extremely powerful as they provide authenticity, integrity, and privacy of communications between two parties.

@@

A smaller key size is one of the most significant benefits introduced by ECC. Current NIST recommendations state that a key size of 160 bits for ECC offers a similar level of security to 1024 bits of RSA/DH. Likewise, 224 bits for ECC is analogous to 2048 bits of RSA/DH. It should be noted that the required key sizes for ECC increases significantly slower than the key sizes for RSA and Diffie-Hellman. Thus ECC offers more security per increase in key size than for RSA. The smaller key size also directly correlates to an increase in computational efficiency; as the bit size increases, the speed difference between Diffie-Hellman and ECC grows superlinearly.\cite{CaseForECC}

@@

Once the SSL/TLS handshake completes, both parties use the generated session key and turn to symmetric-key encryption for hiding of data. Common symmetric-key algorithms include Data Encryption Standard, Advanced Encryption Standard, and RC4.

Both DES and AES are block ciphers, whereas RC4 is a stream cipher. DES, shown to be susceptible to brute-force attacks in 1998, has been replaced by triple-DES, (3DES) which itself has been succeeded by the much-stronger AES cipher. However, 3DES is still used on the web and by older Microsoft products.\cite{NIST2007}\cite{Bogdanov2011}\cite{Nikolic2009} The RC4 cipher is simple and extremely fast in software, and is now the most widely used software stream cipher, most notably in protocols such as TLS and WEP.

Robert Graham's analysis of the incoming handshakes to his exit node revealed that almost 90\% of the circuits agreed on the AES cipher, while the remainder used 3DES. For communication to web servers, RC4 remained the most popular algorithm.\cite{Graham2013}

@@

 Although it is popular, a number of practical attacks and weaknesses have been demonstrated against its Key Scheduling Algorithm and PRNG that call into question RC4's security. In November 2013, Microsoft recommended disabling RC4 from TLS wherever possible.
